% Tables from https://docs.google.com/spreadsheets/d/1DiFTjsC4dP8XyOV7-uF0zwkl0r0jMuW9U9uELejpmn8/edit#gid=0
\section{Introduction}
The sizing and cost model for DM has not been revised since 2014. This document presents
a simplified sizing model in \secref{sec:sizemodel} based on detailed sizing presented in \secref{sec:sizeinputs}.
\secref{sec:cost} presents a very high level budget summary for
DM hardware.

As we begin to consider operations one possibility would be to consider the DR1 hardware
as an operations cost and only consider the commissioning hardware a construction cost.
This is laid out in \secref{sec:scope}.

\section{Proposed budget}\label{sec:cost}

A high level bottom line is given in \tabref{tab:Summary}.
Since Xeon is more expensive that is the number used for the budget calculation, should we get Rome and it really performs we may save a little.
The remainder of the document is all the details that went into that.

\input{Summary}

In \tabref{tab:Summary} we should note that IN2P3 do 50\% of processing so we reduce the processing cost by half. This does not
reduce the storage cost. We have applied a modest cost reduction assuming that processors  and disks get a little cheaper - that
percentage is given in \tabref{tab:Inputs} along with many other parameters. \tabref{tab:Inputs} also contains the number of nodes we assume to need for Qserv.

Specific  costs for storage are detailed in \tabref{tab:Storage} and for compute in \tabref {tab:Machines}
the following budgets can be considered. The detailed annual purchasing based on those prices is given for
storage in \tabref{tab:StorageCost} and for compute in \tabref{tab:Xeon}.

\section{Potential scope option} \label{sec:scope}
In the 2019 JSR we discussed the possibility of purchasing DR1 hardware as part of operations
rather than DM construction.
\tabref{tab:Scope} defines what this would be worth using the cost/sizing model in this document.
The table also calls out explicitly the storage  and compute costs in Chile for the Data Access centre,
these are not deducted but shown to give an idea as to the potential cost there.

\input{Scope}

Should we do this some contingency for extra hardware must be kept in DM construction as well as some manpower
budget to aid with the transition to operations.

\input{ops}

\section{Cost details}
The summary table (\tabref{tab:Summary})  uses Xeon pricing for compute as shown in
\tabref{tab:Xeon}.
\input{Xeon}

An alternative architecture would be Rome - we have not tested this but if it performs well it may provide savings.
\tabref{tab:Rome} gives the price of compute based on Rome -small and large only for comparison purposes - not used in
the calculations.
\input{Rome}

\tabref{tab:StorageCost} gives the price of storage using all  types that we need.
This would be needed regardless of the compute chosen.
\input{StorageCost}

\tabref{tab:overheadCost} gives the annual cost of hosting compute. This includes purchasing racks to house
new nodes etc.
\input{overheadCost}

\input{opsdetails}


\section{ Models}\label{sec:model}
\subsection{Sizing model}\label{sec:sizemodel}

An exhaustive and detailed mode is provided in  \citedsp{LDM-138,LDM-144} - here we concentrate on the needs for the
final years of construction. We explore the compute and storage needed to get us through commissioning and suggest a 2023 purchase for DR1,2 processing which could be pushed to operations.

\tabref{tab:Inputs} gives the annual requirements for the next few years.

\input{Inputs}

\subsection{Compute and storage }\label{sec:csmodel}
We which to base our budget on reasonable well know machines for which we have well know prices.
\tabref{tab:Machines} gives an outline of a few standard machines we use and a price. This table also gives a FLOP estimate
for those machines.
\tabref{tab:Storage} gives costs for different types of storage - we will require various latency for different tasks
and those have varying costs.
These tables are used as look ups for the cost models in \secref{sec:cost}

\input{Storage}

In \tabref{tab:Storage} we should consider for NVME for each TB with file system servers two DDN NVME box with GPFS servers.
The price is based on the TOP performer with best price .
The Normal price is for each TB with file system disks and servers locally attached to production resources.

In the latency and high latency prices are only at NCSA: for each TB with file systems and all people/services.
The complete service not usually attached.   S3 bucket type.
Can be mounted if needed but not for production worthy speeds.
The complete service with data flowing to tape using policies.

\input{Machines}

There is also an associated running cost for machines included in the total cost of ownership.
These overheads are listed in \tabref{tab:overheads}.

\input{overheads}


