\section{Operations budget estimate}\label{sec:opscost}
Based on the needs in \tabref{tab:opsInputs} and the costs in \tabref{tab:opsStorageCost} and \tabref {tab:Machines}
we get the estimate presented in \tabref{tab:opsSummary}.
In \tabref{tab:opsSummary} we should note that IN2P3 do 50\%  and UKDF do 25\% of the processing so we reduce the processing cost by three quarters.
This does not reduce the storage cost.

\input{opsSummary}

Again in \tabref{tab:opsSummary} we assume IN2P3 do 50\% of processing (see \tabref{tab:opsSumUSDF}).
We have applied a compounded modest cost reduction assuming that processors  and disks get a little cheaper - that
percentage is given in \tabref{tab:Inputs}.

It must be noted that the price of disk and tape have a profound effect over 10 years. We have been fairly conservative on the base prices
in \tabref{tab:Storage}. An even bigger effect is in the compounding of the presumed fall in storage cost. Here we have used an extremely
conservative 5\% per year (\tabref{tab:Inputs}) - changing this to 15\% halves the cumulative ops estimate, setting it to 10\% brings the
total down by about 30\%.

\subsection{Cloud costs}
In addition there are some cloud costs. We run certain jobs and host websites on Amazon and Google. In operations
the validation team may also wish to run simulations on cloud resources. This estimate is in \tabref{tab:cloud}.

\input{cloud}


More details on the inputs are in \secref{sec:opsdetails}

\subsection{US and Chile}
While \tabref{tab:opsSummary} present the total ops cost for Rubin Observatory a fraction of this is in Chile and would potentially remain an NSF cost in operations. \tabref{tab:opsSumUSDF} presents just the US Data Facility budget  and
\tabref{tab:opsChileR} presents the Chile budget.

\input{opsSumUSDF}

Note that in \tabref{tab:opsChileR} a fraction of the USDF storage is considered, big enough for Raws and a cache of products. It also does not include a Qserv. See also \tabref{tab:opsSumChile}
\input{opsChileR}
